---
title: "Response by Station"
author: "T. Poole"
date: '2022-05-06'
output: 
  html_document:
          code_fold: hide
editor_options: 
  markdown: 
    wrap: 72
---

## SBFD Response Time Analysis of 5 Residental Occupancies

This is a look at response time for by Ladder 1 and Engine 2 to six
occupancies. Data was queried from our records management system for six
years between 2021 and 2016. Two separate queries where constructed that
resulted in "pivot20162021.csv" and "unitresp20102021.csv". Those files
were joined into a dataframe called "Left_joinedpivotresp.df".

Columns extracted from that dataframe were incident number, address
data, station, apparatus responding and response time in seconds. Those
columns were then filtered for all calls for 90 Allen road, 20
Harborview, 310 Market, 1510/1530 Williston Rd and 465 Quarry Hill,
resulting in 3301 rows of the six variables. A summary of reponses to
each address by Ladder 1 and Engine 2 are broken out below.

*Responses by Engine 1 and Engine 3 are omitted due to contaminated
data.*

```{r,echo=FALSE,include=FALSE}

#### Load Librarys ####

library(tidyverse)

#### Load Data from csv and create data frames#

    ##   Import data and remove non-essential fields/variables ####

Pivot_Table_2016_2021.df <- read.csv(file = "pivot20162021.csv",header = TRUE, na.strings=c("", "NA"),sep = ",")
Unitresp20102021.df <- read.csv(file = "unitresp20102021.csv",header = TRUE, na.strings=c("", "NA"),sep = ",")
        
                ## Left join both df above ##
                
Left_joinedpivotresp.df <- left_join(Pivot_Table_2016_2021.df,Unitresp20102021.df, by = "inci_no")
                        
####   Build new data frame and filter for 465 Quarry Hill, 90 Allen road, 1510-1530 Williston, 20 Harborview, 310 Market ####
        
                Frequent_healthcare.df <- Left_joinedpivotresp.df %>%
                        select(inci_no,station,number,street,resp,unit_desc) %>%
                        filter(number %in% c("90","1510","465","1530","62","118","192"),street %in% c("Williston","Quarry Hill","Quarry Hill Road", "Allen", "Tilley", "Tilley Drive"))
                
```

## 1510 and 1530 Williston Road

These two occupancies represent the first "address" we will look at. The
six number summary below gives an overview of 189 Ladder One responses.
The min. value is zero seconds, signifying that at least once the ladder
was on scene when a second call was received to that address. The first
quartile or 25% of the calls were equal or less than 300 seconds. The
Median represents exactly half way point of the data, 50% of the
responses were above that time and 50% were below. The Mean is the
average time for the counted responses. The "3rd Qu." or third quartile
is where 75% of the calls reside at 7 minutes (420 seconds) or below.

#### Ladder One summary (189 calls)

```{r}
Gazebo_ladder1.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("1510","1530"), street %in% c("Williston"), unit_desc %in% c("Ladder 1"))
                        summary(Gazebo_ladder1.df$resp)
```

#### Engine 2 summary (39 calls)

```{r}
Gazebo_engine2.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("1510","1530"), street %in% c("Williston"), unit_desc %in% c("Engine 2"))
                        summary(Gazebo_engine2.df$resp)
```

#### Visuals of The Data

The red dashed line is NFPA 1710, the dotdashed lines show the median
and the black dashed lines illustrate the mean values.

```{r, figures-Gazebo, fig.show="hold", out.width="50%"}
Gazebo_ladder1.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                ggtitle("Ladder 1 Response to 1510/1530 Williston rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Gazebo_ladder1.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Gazebo_ladder1.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

Gazebo_engine2.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                
                                ggtitle("Engine 2 Response to 1510/1530 Williston rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Gazebo_engine2.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Gazebo_engine2.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

                       
                              
```

## 465 Quarry Hill Road

#### Ladder 1 summary (223 calls)

```{r}
Quarry_ladder1.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("465"), street %in% c("Quarry Hill","Quarry Hill Road"), unit_desc %in% c("Ladder 1"))
                        summary(Quarry_ladder1.df$resp)
```

#### Engine 2 summary (64 calls)

```{r}
Quarry_engine2.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("465"), street %in% c("Quarry Hill","Quarry Hill Road"), unit_desc %in% c("Engine 2"))
                        summary(Quarry_engine2.df$resp)
```

```{r, figures-Quarry, fig.show="hold", out.width="50%"}
Quarry_ladder1.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                ggtitle("Ladder 1 response to 465 Quarry Hill rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Quarry_ladder1.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Quarry_ladder1.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

Quarry_engine2.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                
                                ggtitle("Engine 2 Response to 465 Quarry Hill rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Quarry_engine2.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Quarry_engine2.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

                       
                              
```

## 90 Allen Road

#### Ladder 1 summary (14 calls)

```{r}
Allen_ladder1.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("90"), street %in% c("Allen"), unit_desc %in% c("Ladder 1"))
                        summary(Allen_ladder1.df$resp)
```

#### Engine 2 summary (432 calls)

```{r}
Allen_engine2.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("90"), street %in% c("Allen"), unit_desc %in% c("Engine 2"))
                        summary(Allen_engine2.df$resp)
```

```{r, figures-Allen, fig.show="hold", out.width="50%"}
Allen_ladder1.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                ggtitle("Ladder 1 response to 90 Allen rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Allen_ladder1.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Allen_ladder1.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

Allen_engine2.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                
                                ggtitle("Engine 2 Response to 90 Allen rd.") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Allen_engine2.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Allen_engine2.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

                       
                       
```

```{r}
Tilley_ladder1.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("118","62","192"), street %in% c("Tilley", "Tilley Drive"), unit_desc %in% c("Ladder 1"))
                        summary(Tilley_ladder1.df$resp)
```

```{r}
Tilley_engine2.df <- Frequent_healthcare.df %>% 
                       filter(number %in% c("118","62","192"), street %in% c("Tilley", "Tilley Drive"), unit_desc %in% c("Engine 2"))
                        summary(Tilley_engine2.df$resp)
```

```{r, figures-Tilley, fig.show="hold", out.width="50%"}
Tilley_ladder1.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                ggtitle("Ladder 1 response to Tilley Dr addresses") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Tilley_ladder1.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Tilley_ladder1.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

Tilley_engine2.df %>%
                                ggplot(aes(resp)) +
                                geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.7, bw=30) +
                                
                                ggtitle("Engine 2 Response to Tilley Dr addresses") +
                                geom_vline(xintercept = 300, linetype="dashed", color = "red", size=1) +
                                geom_vline(xintercept = mean(Tilley_engine2.df$resp), linetype ="dashed", size = .5) + 
                                geom_vline(xintercept = median(Tilley_engine2.df$resp), linetype ="dotdash", size = .75) + 
                                scale_y_continuous(labels = NULL) +
                                xlab("Seconds") +
                                ylab("Frequency")

```

# NFPA 1710 2020 edition

### 3.3.44.1 Advanced Life Support (ALS).

Emergency medical treatment beyond basic life support that provides for
advanced airway management including intubation, advanced cardiac
monitoring, defibrillation, establishment and maintenance of intravenous
access, and drug therapy.

### 3.3.44.2 \* Basic Life Support (BLS).

A specific level of pre¬hospital medical care provided by trained
responders, focused on rapidly evaluating a patient's condition;
maintaining a patient's airway, breathing, and circulation; controlling
external bleeding; preventing shock; and preventing further injury or
disability by immobilizing potential spinal or other bone fractures.

### 3.3.64.7 Travel Time.

The time interval that begins when a unit is en route to the emergency
incident and ends when the unit arrives at the scene.

### 3.3.64.8 Turnout Time.

The time interval that begins when the emergency response facilities
(ERFs) and emergency response units (ERUs) notification process begins
by either an audible alarm or visual annunciation or both and ends at
the beginning point of travel time.

### 4.1.2.1 The fire department shall establish the following performance objectives for the first-due response zones that are identified by the AHJ:

1.  Alarm handling time completion in accordance with 4.1.2.3

2.  80 seconds turnout time for fire and special operations response and
    60 seconds turnout time for EMS response

3.  \*240 seconds or less travel time for the arrival of the first
    engine company at a fire suppression incident

4.  360 seconds or less travel time for the arrival of the second
    company with a minimum staffing of 4 personnel at a fire suppression
    incident

5.  For other than high-rise, 480 seconds or less travel time for the
    deployment of an initial full alarm assignment at a fire suppression
    incident

6.  For high-rise, 610 seconds or less travel time for the deployment of
    an initial full alarm assignment at a fire suppression incident

7.  240 seconds or less travel time for the arrival of a unit with first
    responder with automatic external defibrillator (AED) or
    higher-level capability at an emergency medical incident

8.  480 seconds or less travel time for the arrival of an advanced life
    support (ALS) unit at an emergency medical incident, where this
    service is provided by the fire department provided a first
    responder with an AED or basic life support (BLS) unit arrived in
    240 seconds or less travel time
